Rationality, Regret, Evolution \& the Environment

### 2015-11-04 paper/paper.tex
rewrite section 3

\title{Rationality, Evolution \& the Environment}
\author{Paolo Galeazzi \& Michael Franke}
\date{current version: November 2nd 2015}

@@

Evolutionary game theory has become an established philosophical tool for probing into
conceptual issues whose complexity requires mathematical modeling. However, most applications
of EGT look at evolutionary selection of behavior in a single \emph{stage game}. There are
actually two crude but common simplifications in this approach. Firstly, the stage game is
usually considered fixed, closed and immutable. Secondly, what evolves are behavioral
strategies for the stage game.

It has consequently been argued that, instead of focusing on a fixed, immutable stage game and
the behavioral selection for that stage game alone, we should rather study the evolutionary
selection of \emph{choice mechanisms} that are adapted in a rich and variable environment
\citep[e.g.][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}. However,
formal accounts of the adaptive value of choice mechanisms have focused, for the most part, on
the evolutionary benefit of learning rules
\citep[e.g.][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}. \todo{More
  references Harley 1981!!!, O Conor \dots}  Moreover, only little attention has been paid to
modeling the direct (game-theoretic) meta-competition between alternative choice mechanisms
\citep[see][for related criticism]{FawcettHamblin2013:Exposing-the-be}.

This makes two contributions to the literature on evolution of choice mechanisms: one
methodological and one conceptual. Firstly, on the methodological side, we introduce a
conservative extension of standard EGT that allows us to reuse notions such as evolutionary
stability / evolutionary dynamics for the study of the evolutionary competition between choice
mechanisms in a statistically variable environment. Concretely, we look at the evolutionary
competition of choice mechanisms in a ``meta-game'' that consists of the average expected
payoffs of choice mechanisms when playing arbitrary games (from a given class). In this sense,
a meta-game captures (the modeller's) assumptions about the relevant statistics of the
environment in which evolutionary forces operate.

Secondly, and in order to show that our method is useful for a conceptual debate about
evolution of choice mechanisms, we argue that paying attention to the interaction of subjective
belief and subjective preference representations is important. Since almost any reasonable
learning or decision-making process will make use of some (subjective, agent-internal (possibly
unconscious)) representation of the quality of choice options, the question of which such
representations are ecologically valuable (i.e., lead to high fitness) is central. Na\"ive
intuition might convince us that preferences that veridically represent the actual, true
fitness measure that evolutionary selection is based upon must surely be the best. We
demonstrate that, if we take into account how agents represent uncertainty, this na\"ive
conclusion is premature.

@@

It could be objected that, without additional constraints, the notion of rationality is likely
to collapse, as it seems possible to deem rational everything that is observed, given the
freedom to adjust beliefs and preferences at will. But there are many ways in which
ascription of beliefs and preferences is constrained further by normative considerations of
rationality. To begin with, if the DM's uncertainty is about objective chance, it can be argued
that subjective beliefs should conform (at least roughly) to objective chance. When the DM's
uncertainty cannot be anchored in an objective notion of chance, then temporal consistency in
his decision making should at least constrain his subjective beliefs, else he fall prey to
exploitation by Dutch books. Similarly, in certain situations it may make sense to postulate on
normative grounds that subjective preferences should be oriented towards tracking objective
utility, e.g., fitness. For instance, in their classic work \myalert{(Alchian, Friedman)} argue
that profit maximization is a reasonable assumption for characterizing outcomes in a
competitive market because only firms behaving in a manner that is consistent with profit
maximization will survive in the long run.\todo{@Paolo: please add references}

@@

The method by which we want to advance our argument is to integrate ideas from the recent
literature on preference evolution (CITE) with basic tenets of adaptationism: we are interested
in studying the evolutionary competition of choice mechanism, conceived here as pairs of
subjective preference and subjective belief representation. Evolutionary competition between
choice mechanisms is driven by the expected objective fitness that each choice mechanisms
accrues when put to use repeatedly in a range of choice situations. We focus on the more
complex case of multi-agent decision making first, i.e., game situations. Solitary decision
making is then easy to deal with subsequently.

\section{The cast}
\label{sec:basic-notions}

\subsection{Games} \label{sec:games}

@@

With some due simplification, there are two types of approaches to solving games: one focuses
on the individual agents' perpectives and gives \emph{choice rules}, another considers sets of
choices of all agents that are, in some sense or other, locked-in. We will look at the former
here; the latter will be relevant only later, when turn to evolutionary competition between
choice mechanisms in Section~\ref{sec:model}.\footnote{We make a subtle distinction here
  between \emph{choice rules} and \emph{choice mechanisms}: the former refers to a function of
  mapping (subjective) preferences and uncertainty unto choices of acts; the latter is more
  encompassing and contains the choice rule, together with the particular way of forming or
  representing preferences and uncertainty.}

\paragraph{Choice rules.} The classical notion of rationality from
Definition~\ref{def:rationality} also applies to games. Let $A_{-i} = \bigtimes A_j$ for all
$j \neq i$ be the set of all \emph{profiles} of acts for all players except $i$. If player $i$
has a belief $\mu_i \in \Delta(A_{-i})$ about the behavior of his opponents, he chooses
rationally only if his choice $a^*_i$ maximizes expected utility
$\EU(a) = \sum_{\act_{-i} \in \Acts_{-i}} \mu_i(\act_{-i}) \myts \pi_i(a_i, a_{-i})$. So this
is a choice rule, based on \emph{maximization of expected utilities}. It requires a
probabilistic belief about the opponents' action choices. In the example from
Figure~\ref{coordgame1}, a rational player would choose $I$ exactly when he believes that his
opponent will choose $I$ with a probability at least $\nicefrac{2}{3}$.

If an agent does not have a precise probabilistic belief (see also
Section~\ref{sec:impr-prob-beli} below), other rules of choice are available. Perhaps the most
well-known, the \emph{maximin rule}, would tell the agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios. For a strategic game with 2 players this can be
formalized as
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{j} \in \Acts_j} \pi_i(\myact{i},
\myact{j})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible out come for each row act? It is $0$ in either
case. Consequently, the maximin rule would not favor either act; players choosing by the
maximin rule would be indifferent between acts $I$ and $II$.

@@

\subsection{Regret theory} \label{sec:regreTheory}

Regret theory has been introduced in decision theory by \citet{Savage1951:The-theory-of-s} and
Niherens [], and later developed by Bell [1982], Loomes and Sugden [1982], and Fishburn [1982]
independently. Lately, \citet{HalpernPass2012:Iterated-Regret} showed how the use of regret
minimization can give solutions to game theoretical puzzles (like the Traveller's dilemma, see
Example~\ref{example:TravelersDilemma}) in a way that is closer to everyday intuition and
empirical data. In this paper the notion of regret is defined as in
\citet{HalpernPass2012:Iterated-Regret}:

@@

\begin{example}[Coordination Game]
  Let us consider the coordination game in Figure \ref{coordgame1}. Figure~\ref{coordgame1reg}
  shows the pairwise regrets and the regrets of each row-player action. We have
  $\text{reg}(I)=2$, because when the other player chooses action $I$ the pairwise regret
  $reg(I,I)$ is $0$ since $I$ is a best reply to $I$, and when the other player chooses $II$
  the pairwise regret $reg(I, II)$ is $2$ since $\pi_i(II,II)-\pi_i(I,II)=2$. With the same
  reasoning, $\text{reg}(II)=1$. A regret minimizing player would therefore choose act
  $II$. $ \medsquare $

@@

\begin{example}[The Traveller's Dilemma \citep{HalpernPass2012:Iterated-Regret}]
\label{example:TravelersDilemma}
Let us now consider the Traveller's Dilemma case. The story behind the game goes as
follows. There are two travellers, Ann and Bob, who lost their luggages. Ann and Bob had
exactly the same luggage and they had insured the luggages with the same insurance company. The
insurance policy is that the two travellers have to separately claim a certain amount between
2\$ and 100\$ for the reimbursement, and if they both claim the same amount then they will be
reimbursed by that amount. Otherwise, if one of them claims more than the other, then the one
who claimed the higher amount will get the lower amount minus 2\$, while the one who claimed
the lower amount will get what (s)he claimed plus 2\$. The game is symmetric, and the payoff
function for both players is:
\begin{align*}
  \pi_i(a_i,a_{-i}) = \begin{cases} a_i \text{ \hspace{1cm} if } a_i = a_{-i} \\ a_i + 2 \text{
      \hspace{.4cm} if } a_i < a_{-i} \\ a_{-i} -2 \text{ \hspace{.25cm} if } a_{-i} <
    a_i \end{cases}
\end{align*}
By a simple computation we have:  $\forall a_i \in \lbrace 96\$, ..., 100\$ \rbrace \text{, }
\text{reg}(a_i)=3 $ and $\forall a_i \in \lbrace 2\$, ..., 95\$ \rbrace \text{, }
\text{reg}(a_i)>3$. If we now eliminate all the actions that do not minimize regret in the
first place and iterate the regret minimization only on the action set $\lbrace 96\$, ...,
100\$ \rbrace$, we get that 97\$ is the action that minimize regret. Indeed, $\text{reg}(97)=2$
and $\forall a_i \in \lbrace 96\$, 98\$, 99\$, 100\$ \rbrace \text{, } \text{reg}(a_i)=3
$. Hence, one step of regret minimizing reasoning eliminates all the actions smaller than 96\$,
and if we iterate the process once more the unique outcome is 97\$. Finally, it is worth
noticing that the only rationalizable equilibrium of the game corresponds to the outcome
$(2,2)$, that is consequently the only Nash equilibrium, the only perfect equilibrium, and the
only sequential equilibrium of the game.

\begin{itemize}
\item \textcolor{red}{add here:}
  \begin{itemize}  
  \item outcome of rationalizability (max-EU)
  \item outcome of maximin rule
  \item outcome of Laplace rule
  \end{itemize}

 $ \medsquare $

@@

\noindent The acts selected by the maximin rule for the negative regret transformation are exactly
the acts selected by regret minimization for the original game.

This means that we can think of the maximin rule and regret minimization as being two
instantiations of the same choice rule, but operating on different subjective representations
of utility: while maximin considers actual objective payoffs, regret minimization considers
non-veridical, regret-based subjective utilities. As the example of Figure~\ref{coordgame1}
shows, the same choice rule can give different behavioral predictions on different preference
representations. While maximin is indifferent between $I$ and $II$, regret minimization
uniquely selects $II$.

When we think of regret as a non-veridical subjective payoff representation, we can also pair
it with other choice rules, such as the maximization of expected utility and Laplace's rule
(which is just a special case of the former). It turns out, however, that this makes no
noticeable difference:

\noindent In other words, given the same probabilistic belief, maximization of expected
(objective) payoff and minimization of expected (subjective) regret are behaviorally
equivalent. This holds for arbitrary games, as well as for solitary decision situations.

\todo[inline]{Minor edits up to here}

@@

@@

Before concluding this section, it is worth stressing that player $(\tau^{\pi}, \Delta(A_{-i}))$ turns out to be the objective maximinimizer, player $(\tau^{reg}, \Delta(A_{-i}))$ is the regret minimizer, player $(\tau^{\pi}, \overline{\mu})$ is the objective expected utility maximizer, and player $(\tau^{reg}, \overline{\mu})$ is the expected regret minimizer. Moreover, the following fact holds.

@@

Simulation results of the (discrete time) \emph{replicator dynamics}
\citep{TaylorJonker1978:Evolutionary-St} indeed show that random initial population
configurations are attracted to states with only three player types:
$\tuple{\tau^{reg}, \Delta(A)}$, $\tuple{\tau^{reg}, \bar{\mu}}$ and
$\tuple{\tau^{\pi}, \bar{\mu}}$. The relative proportions of these depend on the initial
population.\todo{check uniform notation for player types: round vs.~square brackets} This
variability disappears if we add a small mutation rate to the dynamics. We assume a fixed,
small mutation rate $\epsilon$ for the probability that a player's preference type \emph{or}
her epistemic type changes to another random preference type or epistemic type. The probability
that a player type randomly mutates into a completely different player type with altogether
different preference type and epistemic type would then be $\epsilon^2$. With these assumptions
about ``component-wise mutations'', numerical simulations of the (discrete time)
\emph{replicator mutator dynamics} \citep{Nowak2006:Evolutionary-Dy} show that already for very
small mutation rates almost all initial populations converge to a single fixed point in which
the majority of players are regret types. For instance, with $\epsilon = 0.001$, almost all
initial populations are attracted to a final distribution with proportions:\todo{recalculate
  these numbers for the actual 4x4 game!}

\begin{center}
  \begin{tabular}{ccc}
    $\tuple{\tau^{reg}, \Delta(A)}$ & $\tuple{\tau^{reg},
      \bar{\mu}}$ & $\tuple{\tau^{\pi}, \bar{\mu}}$ \\ \hline
    0.279  &   0.383 &    0.281 
  \end{tabular}
\end{center}

@@

\begin{table}[]
\centering
\footnotesize
\begin{tabular}{ccccccccc}
  \hline
 & $\tuple{\tau^{reg}, \Delta(A)}$ 
 & $\tuple{\tau^{\pi}, \Delta(A)}$ 
 & $\tuple{\tau^{com}, \Delta(A)}$
 & $\tuple{\tau^{alt}, \Delta(A)}$
 & $\tuple{\tau^{reg}, \bar{\mu}}$ 
 & $\tuple{\tau^{\pi}, \bar{\mu}}$ 
 & $\tuple{\tau^{com}, \bar{\mu}}$
 & $\tuple{\tau^{alt}, \bar{\mu}}$ \\ 
  \hline
  $\tuple{\tau^{reg}, \Delta(A)}$ & 6.663 & 6.662 & 5.829 & 7.105 & 6.663 & 6.663 & 5.829 & 7.489 \\
  $\tuple{\tau^{\pi}, \Delta(A)}$ & 6.486 & 6.484 & 6.088 & 6.703 & 6.486 & 6.486 & 6.088 & 6.875 \\
  $\tuple{\tau^{com}, \Delta(A)}$ & 6.323 & 6.758 & 5.496 & 6.977 & 6.323 & 6.323 & 5.496 & 7.149 \\
  $\tuple{\tau^{alt}, \Delta(A)}$ & 5.949 & 5.722 & 5.326 & 6.396 & 5.949 & 5.949 & 5.326 & 6.568 \\
  $\tuple{\tau^{reg}, \bar{\mu}}$ & 6.663 & 6.662 & 5.829 & 7.105 & 6.663 & 6.663 & 5.829 & 7.489 \\
  $\tuple{\tau^{\pi}, \bar{\mu}}$ & 6.663 & 6.662 & 5.829 & 7.105 & 6.663 & 6.663 & 5.829 & 7.489 \\
  $\tuple{\tau^{com}, \bar{\mu}}$ & 6.323 & 6.758 & 5.496 & 6.977 & 6.323 & 6.323 & 5.496 & 7.149 \\
  $\tuple{\tau^{alt}, \bar{\mu}}$ & 6.331 & 5.893 & 5.497 & 6.566 & 6.331 & 6.331 & 5.497 & 7.152 \\
   \hline                          
\end{tabular}                      
\caption{Average payoff in simulations of 100,000
  randomly generated $2 \times 2$ symmetric games.}
\label{tab:ExpectedFitness_2x2_Full}        
\end{table}

>>>>

\title{Rationality, Regret, Evolution \& the Environment}
\author{Paolo Galeazzi \& Michael Franke}
\date{current version: November 4, 2015}

@@

Evolutionary game theory has become an established philosophical tool for probing into
conceptual issues whose complexity requires mathematical modeling. However, most applications
of EGT look at evolutionary selection of behavior in a single \emph{stage game}. There are
actually two crude but common simplifications in this approach. Firstly, the stage game is
usually considered fixed, closed and immutable. Secondly, what evolves are behavioral
strategies for the stage game. It has consequently been argued that, instead of focusing on a
fixed, immutable stage game and the behavioral selection for that stage game alone, we should
rather study the evolutionary selection of \emph{choice mechanisms} that are adapted in a rich
and variable environment
\citep[e.g.][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}. % However,
% formal accounts of the adaptive value of choice mechanisms have focused, for the most part, on
% the evolutionary benefit of learning rules
% \citep[e.g.][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}. \todo{More
%   references Harley 1981!!!, O Conor \dots} Moreover,
But only little attention has been paid to modeling the evolutionary competition between choice
mechanisms in a game-theoretic framework \citep[but
see][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}.\todo{More references
  Harley 1981!!!, O Conor \dots}.

This paper makes two contributions to the literature on evolution of choice mechanisms: one
methodological and one conceptual. Firstly, on the methodological side, we introduce a
conservative extension of standard EGT that allows us to reuse notions such as evolutionary
stability / evolutionary dynamics for the study of the evolutionary competition between choice
mechanisms in a statistically variable environment. Concretely, we look at the evolutionary
competition of choice mechanisms in a ``meta-game'' whose payoffs represent the average
expected fitness choice mechanisms when playing arbitrary games (from a given class) against
themselves. In this sense, a meta-game captures (the modeller's) assumptions about the relevant
statistics of the environment in which evolutionary forces operate.

Secondly, and in order to show that our method is useful for a conceptual debate about
evolution of choice mechanisms, we argue that paying attention to the interaction of subjective
belief and subjective preference representations is important. Since almost any reasonable
learning or decision-making process will make use of some (subjective, agent-internal (possibly
unconscious)) representation of the quality of choice options, the question of which such
representations are ecologically valuable (i.e., lead to high fitness) is central. Na\"ive
intuition might convince us that preferences that veridically represent the actual, true
fitness measure that evolutionary selection is based upon must surely be the best. We
demonstrate that, if we take into account how agents represent uncertainty, this na\"ive
conclusion is premature. Under certain reasonable epistemic conditions, non-veridical
preference representations in terms of regret can be evolutionarily beneficial.

@@

It could be objected that, without additional constraints, the notion of rationality is likely
to collapse, as it seems possible to deem rational everything that is observed, given the
freedom to adjust beliefs and preferences at will. But there are many ways in which ascriptions
of beliefs and preferences are constrained by normative considerations of rationality. To begin
with, if the DM's uncertainty is about objective chance, it can be argued that subjective
beliefs should conform (at least roughly) to objective chance. When the DM's uncertainty cannot
be anchored in an objective notion of chance, then temporal consistency in his decision making
should at least constrain his subjective beliefs, else he fall prey to exploitation by Dutch
books. Similarly, in certain situations it may make sense to postulate on normative grounds
that subjective preferences should be oriented towards tracking objective utility, e.g.,
fitness. For instance, in their classic work \myalert{(Alchian, Friedman)} argue that profit
maximization is a reasonable assumption for characterizing outcomes in a competitive market
because only firms behaving in a manner that is consistent with profit maximization will
survive in the long run.\todo{@Paolo: please add references}

@@

The method by which we want to advance our argument is to integrate ideas from the recent
literature on preference evolution (CITE) with basic tenets of adaptationism: we are interested
in studying the evolutionary competition of choice mechanisms, conceived here as pairs of
subjective preference and subjective belief representations. Evolutionary competition between
choice mechanisms is driven by the expected objective fitness that each choice mechanisms
accrues when put to use repeatedly in a range of choice situations. We focus on the more
complex case of multi-agent decision making first, i.e., game situations. Solitary decision
making is then easy to deal with subsequently. The next section introduces the choice
mechanisms we focus on. After that, Section~\ref{sec:model} introduces the ``meta-game''
modeling approach.

\section{Choice mechanisms: preferences, beliefs and action choice}
\label{sec:basic-notions}

@@

With some due simplification, there are two types of approaches to solving games. One looks at
a collection of players (e.g., a whole population of them) and considers sets of choices of all
players that are, in some sense or other, stable or in equilibrium. Another approach is to
focus on each individual agent's perpective and specify which of her actions the agent
\emph{should} choose if she is rational. We will look at the latter here; the former will be
relevant only later, when turn to evolutionary competition between choice mechanisms in
Section~\ref{sec:model}. We will first look at what we call \emph{choice rules}: classical
proposals of how rational agents should choose acts. We will then introduce our notion of
\emph{choice mechanism} which makes explicit several elements that choice rules might brush
under the table.

\subsection{Choice rules} 
\label{sec:choice-rules}

For the purposes of this paper, we think of a choice rule as a function that takes a
(strategic) game as input and delivers a choice of action for that game. The literatures has
suggested a number of choice rules that aspire to capture a normative element of rational
choice. For reasons that will become clear as we go along, we focus here on three classical
choice rules: (i) maximin, (ii) Laplace and (iii) regret minimization.

But first an important digression. Rational choice from maximization of expected utilities, as
given in Definition~\ref{def:rationality}, is \emph{not} a choice rule in the present sense, if
we apply it strictly. Although the notion applies to game situations, of course, it requires an
additional component, namely a belief of the DM, before it can say which acts the DM should
choose. Let $A_{-i} = \bigtimes A_j$ for all $j \neq i$ be the set of all \emph{profiles} of
acts for all players except $i$. If player $i$ has a belief $\mu_i \in \Delta(A_{-i})$ about
the behavior of his opponents, he chooses rationally only if his choice $a^*_i$ maximizes
expected utility
$\EU(a) = \sum_{\act_{-i} \in \Acts_{-i}} \mu_i(\act_{-i}) \myts \pi_i(a_i, a_{-i})$. So this
requires a probabilistic belief about the opponents' action choices. In the example from
Figure~\ref{coordgame1}, a rational player would choose act $I$ exactly when he believes that
his opponent will choose act $I$ with a probability at least $\nicefrac{2}{3}$.

Perhaps the most well-known choice rule that does not require explicitly stating the DM's
beliefs is the \emph{maximin rule}. It tells our agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios. For a strategic game with 2 players this can be
formalized as
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{j} \in \Acts_j} \pi_i(\myact{i},
\myact{j})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible outcome for each row act? It is $0$ in either
case. Consequently, the maximin rule would not favor either act; players choosing by the
maximin rule would be indifferent between acts $I$ and $II$.

@@

The final choice rule that is relevant for our purposes here is \emph{regret
  minimization}. Regret theory has been introduced in decision theory by
\citet{Savage1951:The-theory-of-s} and Niherens [], and later developed by Bell [1982], Loomes
and Sugden [1982], and Fishburn [1982] independently. Lately,
\citet{HalpernPass2012:Iterated-Regret} showed how the use of regret minimization can give
solutions to game theoretical puzzles (like the Traveller's dilemma) in a way that is closer to
everyday intuition and empirical data. In this paper the notion of regret is defined as in
\citet{HalpernPass2012:Iterated-Regret}:

@@

To illustrate, Figure~\ref{coordgame1reg} shows the pairwise regrets and the regrets of each
row-player action for the coordination game in Figure \ref{coordgame1}. We have
$\text{reg}(I)=2$, because when the other player chooses action $I$ the pairwise regret
$reg(I,I)$ is $0$ since $I$ is a best reply to $I$, and when the other player chooses $II$ the
pairwise regret $reg(I, II)$ is $2$ since $\pi_i(II,II)-\pi_i(I,II)=2$. With the same
reasoning, $\text{reg}(II)=1$. A regret minimizing player would therefore choose act $II$.

@@

% \begin{example}[The Traveller's Dilemma \citep{HalpernPass2012:Iterated-Regret}]
% \label{example:TravelersDilemma}
% Let us now consider the Traveller's Dilemma case. The story behind the game goes as
% follows. There are two travellers, Ann and Bob, who lost their luggages. Ann and Bob had
% exactly the same luggage and they had insured the luggages with the same insurance company. The
% insurance policy is that the two travellers have to separately claim a certain amount between
% 2\$ and 100\$ for the reimbursement, and if they both claim the same amount then they will be
% reimbursed by that amount. Otherwise, if one of them claims more than the other, then the one
% who claimed the higher amount will get the lower amount minus 2\$, while the one who claimed
% the lower amount will get what (s)he claimed plus 2\$. The game is symmetric, and the payoff
% function for both players is:
% \begin{align*}
%   \pi_i(a_i,a_{-i}) = \begin{cases} a_i \text{ \hspace{1cm} if } a_i = a_{-i} \\ a_i + 2 \text{
%       \hspace{.4cm} if } a_i < a_{-i} \\ a_{-i} -2 \text{ \hspace{.25cm} if } a_{-i} <
%     a_i \end{cases}
% \end{align*}
% By a simple computation we have:  $\forall a_i \in \lbrace 96\$, ..., 100\$ \rbrace \text{, }
% \text{reg}(a_i)=3 $ and $\forall a_i \in \lbrace 2\$, ..., 95\$ \rbrace \text{, }
% \text{reg}(a_i)>3$. If we now eliminate all the actions that do not minimize regret in the
% first place and iterate the regret minimization only on the action set $\lbrace 96\$, ...,
% 100\$ \rbrace$, we get that 97\$ is the action that minimize regret. Indeed, $\text{reg}(97)=2$
% and $\forall a_i \in \lbrace 96\$, 98\$, 99\$, 100\$ \rbrace \text{, } \text{reg}(a_i)=3
% $. Hence, one step of regret minimizing reasoning eliminates all the actions smaller than 96\$,
% and if we iterate the process once more the unique outcome is 97\$. Finally, it is worth
% noticing that the only rationalizable equilibrium of the game corresponds to the outcome
% $(2,2)$, that is consequently the only Nash equilibrium, the only perfect equilibrium, and the
% only sequential equilibrium of the game.

% \begin{itemize}
% \item \textcolor{red}{add here:}
%   \begin{itemize}  
%   \item outcome of rationalizability (max-EU)
%   \item outcome of maximin rule
%   \item outcome of Laplace rule
%   \end{itemize}

% \end{itemize}

%  $ \medsquare $

% \end{example}

@@

\noindent Acts selected by the maximin rule for the negative regret transformation of $G$ are
always exactly the acts selected by regret minimization for the original game $G$.\todo{can we
  cite a paper that proves this? should we give a proof ourselves?}

This means that we can think of the maximin rule and regret minimization as essentially the
same operation on different subjective representations of preference: while maximin considers
actual objective payoffs (as defined by the modeller's game), regret minimization considers
non-veridical, regret-based subjective preferences. As the example of Figure~\ref{coordgame1}
shows, different subjective preferences can give rise to different choice prescriptions all
else equal. While maximin is indifferent between $I$ and $II$, regret minimization uniquely
selects $II$.

In sum, a choice rule is a normative prescription of action choices for a rational agent in any
game situation. Choice rules may rest on implicit assumptions about what the DM believes (e.g.,
when interpreting Laplace rule as rational choice under flat beliefs). They may also make
implicit assumptions about what the DM prefers (e.g., when reconstructing regret minimization
in terms of maximin-choices on the regret transformation). This is rather messy. To disentangle
and avoid confusion, we turn to choice mechanisms.

\subsection{Choice mechanisms}
\label{sec:choice-mechanisms}

We think of a choice mechanism as a function that maps a game (or more generally any decision
situation) onto a triple that consists of: (i) the DM's \emph{subjective preferences} (possibly
diverging from the actual objective payoffs defined ) in the game, (ii) the DM's
\emph{subjective beliefs} about the opponents' behavior, and (iii) an \emph{action selection
  function} which maps the former two components onto a concrete choice of action in the
game. A choice mechanism could be a set of distinct heuristics (the DM's \emph{adaptive
  toolbox}) that have little in common
\citep[e.g.][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to keep matters simple, we will here only look at choice mechanisms that use a uniform way of
forming preferences and beliefs and a uniform way of making choices based on them. More
concretely, we focus on two \emph{preference types} and two \emph{belief types}, and neglect
variation in the action selection component.

We are most interested here in studying whether evolutionary pressure could support
non-veridical subjective preferences under certain conditions. We therefore focus on two
\emph{preference types}: veridical types $\tau^\pi$, who use the actual payoff function $\pi$ of the
actual game $G$, and regret types $\tau^{reg}$, who look at the regret transformation of the
actual game $G$.

It turns out that whether $\tau^{reg}$ can have a fitness advantage over $\tau^\pi$ depends on
the general \emph{kind} of (probabilistic) beliefs that DM's form about the opponent's
behavior. In other words, we even abstract away from the content of the DM's beliefs about his
opponents' behavior that he could have acquired by learning or observation, but first turn on
the more general question of how uncertainty should be represented in order to interact with
subjective preferences in such a way that the resulting behavior will be fitness enhancing for
the DM in the long run. We consider therefore two \emph{belief types} (to be defined below):
imprecise belief types $\Delta(A_{-i})$, who have a set of probabilistic beliefs about the
opponents' behavior and precise belief types $\overline{\mu}$ who have a single probabilistic
belief.

The following subsection describes what these belief types are and how a single action
selection function (maximin expected utility) gives rise to an intersting parallel with
classical choice rules.

@@

\subsection{Choice mechanisms \& classical choice rules}

Fixing the action selection component to maximin expected utility, and allowing two belief and
two preference types, we end up with four basic choice mechanisms summed up in
Table~\ref{tab:CMs}. Different choice mechanisms can give rise to identical action predictions,
either in a restricted class of games, or in general. This is also the case here.

\begin{table}[t]
  \centering
  \begin{tabular}{ccc}
    preference type & belief type & corresponding choice rule \\ \midrule
    $\tau^\pi$ & $\overline{\mu}$ & Laplace \\
    $\tau^\pi$ & $\Delta(A_{-i})$ & maximin \\ 
    $\tau^{reg}$ & $\overline{\mu}$ & Laplace \\
    $\tau^{reg}$ & $\Delta(A_{-i})$ & regret minimization \\ 
  \end{tabular}
  \caption{Overview relevant choice mechanisms and the corresponding (behaviorally equivalent) classical choice rule}
  \label{tab:CMs}
\end{table}

\noindent In other words, given the same probabilistic belief, maximization of expected
(objective) payoff and minimization of expected (subjective) regret are behaviorally
equivalent. This holds for arbitrary games, as well as for solitary decision situations. On the
other hand, the following does not hold in general:

\begin{fact} \label{fact:equivalence2x2}
In the class of $2 \times 2$ symmetric games, players $(\tau^{reg}, \Delta(A_{-i}))$, $(\tau^{\pi}, \overline{\mu})$, and $(\tau^{reg},\overline{\mu})$ are behaviorally equivalent.
\end{fact}

Moreover, our choice mechanism $(\tau^{\pi}, \Delta(A_{-i}))$ turns out to be the objective
maximinimizer, player $(\tau^{reg}, \Delta(A_{-i}))$ is the regret minimizer, player
$(\tau^{\pi}, \overline{\mu})$ is the objective expected utility maximizer, and player
$(\tau^{reg}, \overline{\mu})$ is the expected regret minimizer. In other words, behaviorally,
the choice mechanisms we defined here correspond to the classical choice rules in
Table~\ref{tab:CMs}.

\bigskip
\bigskip

\todo[inline]{Micha rewrite up to here!!! Please check which bits of text/information from the
  following need to be integrated and what can go.}

\bigskip
\bigskip

@@

@@

Simulation results of the (discrete time) \emph{replicator dynamics}
\citep{TaylorJonker1978:Evolutionary-St} indeed show that random initial population
configurations are attracted to states with only three player types:
$\tuple{\tau^{reg}, \Delta(A)}$, $\tuple{\tau^{reg}, \overline{\mu}}$ and
$\tuple{\tau^{\pi}, \overline{\mu}}$. The relative proportions of these depend on the initial
population.\todo{check uniform notation for player types: round vs.~square brackets} This
variability disappears if we add a small mutation rate to the dynamics. We assume a fixed,
small mutation rate $\epsilon$ for the probability that a player's preference type \emph{or}
her epistemic type changes to another random preference type or epistemic type. The probability
that a player type randomly mutates into a completely different player type with altogether
different preference type and epistemic type would then be $\epsilon^2$. With these assumptions
about ``component-wise mutations'', numerical simulations of the (discrete time)
\emph{replicator mutator dynamics} \citep{Nowak2006:Evolutionary-Dy} show that already for very
small mutation rates almost all initial populations converge to a single fixed point in which
the majority of players are regret types. For instance, with $\epsilon = 0.001$, almost all
initial populations are attracted to a final distribution with proportions:\todo{recalculate
  these numbers for the actual 4x4 game!}

\begin{center}
  \begin{tabular}{ccc}
    $\tuple{\tau^{reg}, \Delta(A)}$ & $\tuple{\tau^{reg},
      \overline{\mu}}$ & $\tuple{\tau^{\pi}, \overline{\mu}}$ \\ \hline
    0.279  &   0.383 &    0.281 
  \end{tabular}
\end{center}

@@

\begin{table}[]
\centering
\footnotesize
\begin{tabular}{ccccccccc}
  \hline
 & $\tuple{\tau^{reg}, \Delta(A)}$ 
 & $\tuple{\tau^{\pi}, \Delta(A)}$ 
 & $\tuple{\tau^{com}, \Delta(A)}$
 & $\tuple{\tau^{alt}, \Delta(A)}$
 & $\tuple{\tau^{reg}, \overline{\mu}}$ 
 & $\tuple{\tau^{\pi}, \overline{\mu}}$ 
 & $\tuple{\tau^{com}, \overline{\mu}}$
 & $\tuple{\tau^{alt}, \overline{\mu}}$ \\ 
  \hline
  $\tuple{\tau^{reg}, \Delta(A)}$ & 6.663 & 6.662 & 5.829 & 7.105 & 6.663 & 6.663 & 5.829 & 7.489 \\
  $\tuple{\tau^{\pi}, \Delta(A)}$ & 6.486 & 6.484 & 6.088 & 6.703 & 6.486 & 6.486 & 6.088 & 6.875 \\
  $\tuple{\tau^{com}, \Delta(A)}$ & 6.323 & 6.758 & 5.496 & 6.977 & 6.323 & 6.323 & 5.496 & 7.149 \\
  $\tuple{\tau^{alt}, \Delta(A)}$ & 5.949 & 5.722 & 5.326 & 6.396 & 5.949 & 5.949 & 5.326 & 6.568 \\
  $\tuple{\tau^{reg}, \overline{\mu}}$ & 6.663 & 6.662 & 5.829 & 7.105 & 6.663 & 6.663 & 5.829 & 7.489 \\
  $\tuple{\tau^{\pi}, \overline{\mu}}$ & 6.663 & 6.662 & 5.829 & 7.105 & 6.663 & 6.663 & 5.829 & 7.489 \\
  $\tuple{\tau^{com}, \overline{\mu}}$ & 6.323 & 6.758 & 5.496 & 6.977 & 6.323 & 6.323 & 5.496 & 7.149 \\
  $\tuple{\tau^{alt}, \overline{\mu}}$ & 6.331 & 5.893 & 5.497 & 6.566 & 6.331 & 6.331 & 5.497 & 7.152 \\
   \hline                          
\end{tabular}                      
\caption{Average payoff in simulations of 100,000
  randomly generated $2 \times 2$ symmetric games.}
\label{tab:ExpectedFitness_2x2_Full}        
\end{table}
